### 2016-09-28 rebuttal.tex

  The major advantage is that mpmath is a pure Python library. It's slower
  than MPFR, but not too much to evaluate any reasonably complicated single
  expression at interactive speed, which covers most use cases in SymPy. We
  have added a sentence about this to the numerics section.

We mention this at the beginning of the section. Another advantage is that
mpmath is BSD licensed, like SymPy, which is important for the scientific
Python ecosystem (MPFR is licensed under the LGPL).
\end{solution}
\question Section 4.1: please state whether the syntax for
functions in mpmath is identical, or not, to that in Sympy.
\begin{solution}
The mpmath library is not a SymPy submodule, but a separate library,
which we have indicated in the text. Most function names in mpmath are the same as SymPy, but it is not true in every case. We have
added an example to the text showing an instance where they diverge
(summations).
\end{solution}
\question Unless I have
missed something, a weakness of the paper is that speed is mentioned only
twice, on lines 496 and 666. My assumption is that SymPy is slow compared with
its commercial competitors. Please comment on the speed issue, and not just in
the conclusions.
\begin{solution}
A new paragraph titled ``Speed'' was added to the manuscript, which discusses the
speed issue and lays out a path forward. The conclusion was simplified.
\end{solution}
\question Editors are needed for [20].
\begin{solution}
We have done this.
\end{solution}
\end{questions}

\section{Reviewer 1}
\subsection{Basic reporting}
\begin{questions}
\question Generally good. A part that confused me is the assertion (footnote 3) that ``If
$A$ and $B$ are Symbols created with \texttt{commutative=False} then SymPy will keep
$A\cdot B$
and $B\cdot A$ distinct.'' Does that mean that BOTH of them must be created this way,
and that $A$ and $x$ (if $x$ is created normally) will commute? Is there any way to
declare commutators? How does one guarantee that other pieces of code, e.g.
Gaussian elimination, respect non-commutativity?
\begin{solution}
  We have clarified the footnote. Both expressions must be set as
  \texttt{commutative=False}. Internally, in \texttt{Mul}, the ``commutative
  part'' of an expression is pulled out to the front and canonically ordered,
  and the ``noncommutative part'' is not reordered.

### 2016-09-28 architecture.tex
new performance

\subsection{Speed}

Due to being written in pure Python, SymPy's speed is generally slower compared
with its commercial competitors. For many applications and uses of SymPy, that
is not a problem, as SymPy is able to return the answer quickly enough, but for
some applications that require handling of very long expressions and/or lots of
small expressions, the speed becomes a problem.

For this reason, a new library called SymEngine~\cite{SymEngine} was started.
It is a pure C++ library with thin wrappers to other languages (Python, Ruby,
Julia, \ldots) whose aim is to be the fastest manipulation library.  Preliminary
benchmarks suggest that SymEngine is as fast or faster than the commercial or
open source competitors.

The development branch of SymPy recently started to use SymEngine as an
optional backend, initially in \texttt{sympy.physics.mechanics} only. The plan is
to allow more algorithms in SymPy to take advantage of the speed of SymEngine.

>>>>

\subsection{Performance}

Due to being written in pure Python without the use of extension modules,
SymPy's performance characteristics are generally poorer than
its commercial competitors. For many applications,
the performance of SymPy, as measured by clock cycles, memory usage, and memory
layout, is sufficient.
However, the boundaries for when SymPy's pure Python strategy becomes
insufficient are when the user requires handling of very long expressions or many
small expressions. Where this boundray lies depends on the system at hand, but tends
to be within the range of $O(10^4) - O(10^6)$ symbols for modern computers.

For this reason, a new project called SymEngine~\cite{SymEngine} has been started.
The aim of this poject is to develop a library with better performance
characteristics for symbolic manipulation. SymEngine is a pure C++ library,
which allows it fine grain control over the memory layout of expressions.
SymEngine has thin wrappers to other languages (Python, Ruby,
Julia, etc.) whose aim is to be the fastest manipulation library. Preliminary
benchmarks suggest that SymEngine is as performant as its commercial and
open source competitors.

SymPy has recently started to use SymEngine as an
optional backend, initially in \texttt{sympy.physics.mechanics} only.
Future work witll involve
allowing more algorithms in SymPy to use SymEngine as a backend.
